---
title: "TEAK Carbon Statistical Analysis"
author: "Marissa Goodwin"
date: "December 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Set Working Directory and Load Library
```{r}
setwd("C:/Users/mjgoodwin/Desktop/TEAK Carbon Manuscript Code")

library("nlme")

library("multcomp")

library(multcompView)

library(rcompanion)

library(ggplot2)

library(lsmeans)

library(lme4)

library(PMCMR)

library(car)

library(FSA)

```


###Repeated Measures ANOVA for Carbon Stocks
```{r pressure, echo=FALSE}
carb <- read.csv("plotaverages.csv")
carb$year <- as.factor(carb$year)

live <- carb[which(carb$fuel_type == "Live"),]
snag <- carb[which(carb$fuel_type == "Dead"),]
cwd <- carb[which(carb$fuel_type == "CWD"),]
fwd <- carb[which(carb$fuel_type == "FWD"),]
lit <- carb[which(carb$fuel_type == "Litter"),]
```

##Live Trees
```{r}
#Checking Assumptions of a Repeated Measures ANOVA
by(live$carbon, live$year, shapiro.test)
#Results show that each year meets the normality assumption

#Check for outliers
boxplot(carbon ~ year, data=live, main="Live Carbon Outliers by Year")
#Results show no outliers for each year

#Check Homogeneity of Variance
bartlett.test(carbon ~ year, live)
#Results indicate failure to reject null hypothesis that variance is the same for all groups

#Repeated Measures ANOVA on Live Carbon by Treatment across Years
live_lme <- lme(carbon ~ Treatment * year, random= ~ year|Plot, data=live, method="ML", control=list(opt="optim"))
#summary(live_lme)
anova(live_lme)
posthoc_live <- glht(live_lme, linfct=mcp(Treatment = "Tukey"))
summary(posthoc_live)
cld(posthoc_live, alpha=0.05, Letters=letters)

#Post Hoc Test Looking at Treatment Differences by Year
posthoc_live_by_year <- glht(live_lme, lsm(pairwise ~ Treatment|year))
summary(posthoc_live_by_year)
```


##Dead Trees
```{r}
#Checking Assumptions of a Repeated Measures ANOVA
by(snag$carbon, snag$year, shapiro.test)
#Results show that each year meets the normality assumption

#Check for outliers
boxplot(carbon ~ year, data=snag, main="Snag Carbon Outliers by Year")
#Results show outliers

#Check Homogeneity of Variance
bartlett.test(carbon ~ year, snag)
#Results indicate to reject null hypothesis that variance is the same for all groups

#Transform Data to meet Assumptions
snag$transcarb <- log(snag$carbon)

#Checking Assumptions of Transfromed Data
by(snag$transcarb, snag$year, shapiro.test)
#Results show that each year meets the normality assumption

#Check for outliers
boxplot(transcarb ~ year, data=snag, main="Snag Carbon Outliers by Year")
#Results show no significant outliers

#Check Homogeneity of Variance
bartlett.test(transcarb ~ year, snag)
#Results indicate failure to reject null hypothesis that variance is the same for all groups

#Repeated Measures ANOVA on Snag Carbon by Treatment across Years
snag_lme <- lme(transcarb ~ Treatment * year, random= ~ year|Plot, data=snag, method="ML", control=list(opt="optim"))
#summary(snag_lme)
anova(snag_lme)
posthoc_snag <- glht(snag_lme, linfct=mcp(Treatment = "Tukey"))
summary(posthoc_snag)
cld(posthoc_snag, alpha=0.05, Letters=letters)

#Post Hoc Test Looking at Treatment Differences by Year
posthoc_snag_by_year <- glht(snag_lme, lsm(pairwise ~ Treatment|year))
summary(posthoc_snag_by_year)
```


##CWD
```{r}
#Checking Assumptions of a Repeated Measures ANOVA
by(cwd$carbon, cwd$year, shapiro.test)
#Results show that not all years meet normality assumption

#Check for outliers
boxplot(carbon ~ year, data=cwd, main="CWD Carbon Outliers by Year")
#Results show outliers

#Check Homogeneity of Variance
bartlett.test(carbon ~ year, cwd)
#Results indicate rejection of null hypothesis that variance is the same for all groups

#Transform Data to meet Assumptions
cwd$transcarb <- log10(cwd$carbon)

#Checking Assumptions of Transfromed Data
by(cwd$transcarb, cwd$year, shapiro.test)
#Results show that each year meets the normality assumption

#Check for outliers
boxplot(transcarb ~ year, data=cwd, main="CWD Carbon Outliers by Year")
#Results show no significant outliers

#Check Homogeneity of Variance
bartlett.test(transcarb ~ year, cwd)
#Results indicate failure to reject null hypothesis that variance is the same for all groups

#Repeated Measures ANOVA on Cwd Carbon by Treatment across Years
cwd_lme <- lme(transcarb ~ Treatment * year, random= ~ year|Plot, data=cwd, method="ML", control=list(opt="optim"))
#summary(cwd_lme)
anova(cwd_lme)
posthoc_cwd <- glht(cwd_lme, linfct=mcp(Treatment = "Tukey"))
summary(posthoc_cwd)
cld(posthoc_cwd, alpha=0.05, Letters=letters)

#Post Hoc Test Looking at Treatment Differences by Year
posthoc_cwd_by_year <- glht(cwd_lme, lsm(pairwise ~ Treatment|year))
summary(posthoc_cwd_by_year)
```

##FWD
```{r}
#Checking Assumptions of a Repeated Measures ANOVA
by(fwd$carbon, fwd$year, shapiro.test)
#Normality Assumption Met

#Check for outliers
boxplot(carbon ~ year, data=fwd, main="FWD Carbon Outliers by Year")
#No Outliers

#Check Homogeneity of Variance
bartlett.test(carbon ~ year, fwd)
#Results indicate rejection of null hypothesis that variance is the same for all groups

#Transform Data to meet Assumptions
fwd$transcarb <- log10(fwd$carbon)

#Checking Assumptions of Transfromed Data
by(fwd$transcarb, fwd$year, shapiro.test)
#Results show that each year meets the normality assumption

#Check for outliers
boxplot(transcarb ~ year, data=fwd, main="FWD Carbon Outliers by Year")
#Results show no significant outliers

#Check Homogeneity of Variance
bartlett.test(transcarb ~ year, fwd)
#Results indicate failure to reject null hypothesis that variance is the same in all groups

#Repeated Measures ANOVA on Fwd Carbon by Treatment across Years
fwd_lme <- lme(transcarb ~ Treatment * year, random= ~ year|Plot, data=fwd, method="ML", control=list(opt="optim"))
#summary(fwd_lme)
anova(fwd_lme)
posthoc_fwd <- glht(fwd_lme, linfct=mcp(Treatment = "Tukey"))
summary(posthoc_fwd)
cld(posthoc_fwd, alpha=0.05, Letters=letters)

#Post Hoc Test Looking at Treatment Differences by Year
posthoc_fwd_by_year <- glht(fwd_lme, lsm(pairwise ~ Treatment|year))
summary(posthoc_fwd_by_year)
```


##Litter
```{r}
#Checking Assumptions of a Repeated Measures ANOVA
by(lit$carbon, lit$year, shapiro.test)
#Results show that all years meet assumption of normality

#Check for outliers
boxplot(carbon ~ year, data=lit, main="Litter Outliers by Year")
#Results show no outliers

#Check Homogeneity of Variance
bartlett.test(carbon ~ year, lit)
#Results indicate failure to reject null hypothesis that variance is the same for all groups


#Repeated Measures ANOVA on Lit Carbon by Treatment across Years
lit_lme <- lme(carbon ~ Treatment * year, random= ~ year|Plot, data=lit, method="ML", control=list(opt="optim"))
#summary(lit_lme)
anova(lit_lme)
posthoc_lit <- glht(lit_lme, linfct=mcp(Treatment = "Tukey"))
summary(posthoc_lit)
cld(posthoc_lit, alpha=0.05, Letters=letters)

#Post Hoc Test Looking at Treatment Differences by Year
posthoc_lit_by_year <- glht(lit_lme, lsm(pairwise ~ Treatment|year))
summary(posthoc_lit_by_year)
```


##Total Carbon
```{r}
tot <- aggregate(carbon ~ Plot + Treatment + year, carb, sum)

#Checking Assumptions of a Repeated Measures ANOVA
by(tot$carbon, tot$year, shapiro.test)
#Results show that all years meet assumption of normality

#Check for outliers
boxplot(carbon ~ year, data=tot, main="Total Carbon Outliers by Year")
#Results show no outliers

#Check Homogeneity of Variance
bartlett.test(carbon ~ year, tot)
#Results indicate failure to reject null hypothesis that variance is the same for all groups

#Repeated Measures ANOVA on Lit Carbon by Treatment across Years
tot_lme <- lme(carbon ~ Treatment * year, random= ~ year|Plot, data=tot, method="ML", control=list(opt="optim"))
#summary(tot_lme)
anova(tot_lme)
posthoc_tot <- glht(tot_lme, linfct=mcp(Treatment = "Tukey"))
summary(posthoc_tot)
cld(posthoc_tot, alpha=0.05, Letters=letters)

#Post Hoc Test Looking at Treatment Differences by Year
posthoc_tot_by_year <- glht(tot_lme, lsm(pairwise ~ Treatment|year))
summary(posthoc_tot_by_year)
```

#Analysis of Variance of Changes in Carbon Stocks
```{r}
diff <- read.csv("carbondiff.csv")

#Drought Period
drought <- diff[(diff$period == "drought"),]

#Checking Normality Assumption for each fuel type (CWD is not normally distributed)
by(drought$diff, drought$fuel_type, shapiro.test)


#Live
live <- drought[(drought$fuel_type == "Live"),]

bartlett.test(diff ~ Treatment, live) # Checking Homogeneity of Variance (checks out)

live_aov <- aov(diff ~ Treatment, data=live)
summary(live_aov)

TukeyHSD(live_aov)

#Dead
dead <- drought[(drought$fuel_type == "Dead"),]

bartlett.test(diff ~ Treatment, dead) # Checking Homogeneity of Variance (checks out)

dead_aov <- aov(diff ~ Treatment, data=dead)
summary(dead_aov)

TukeyHSD(dead_aov)


#CWD
cwd <- drought[(drought$fuel_type == "CWD"),]

cwd$transdiff <- sqrt(cwd$diff+100)

shapiro.test(cwd$transdiff)

bartlett.test(transdiff ~ Treatment, cwd) # Checking Homogeneity of Variance (checks out)

cwd_aov <- aov(diff ~ Treatment, data=cwd)
summary(cwd_aov)

TukeyHSD(cwd_aov)


#FWD
fwd <- drought[(drought$fuel_type == "FWD"),]

bartlett.test(diff ~ Treatment, fwd) # Checking Homogeneity of Variance (checks out)

fwd_aov <- aov(diff ~ Treatment, data=fwd)
summary(fwd_aov)

TukeyHSD(fwd_aov)

#Litter
litter <- drought[(drought$fuel_type == "Litter"),]

bartlett.test(diff ~ Treatment, litter) # Checking Homogeneity of Variance (checks out)

litter_aov <- aov(diff ~ Treatment, data=litter)
summary(litter_aov)

TukeyHSD(litter_aov)


#Reburn Period(looking at only burn plots)
reburn <- diff[(diff$period == "reburn"),]

reburn <- reburn[(reburn$Treatment == "bs"|reburn$Treatment == "bc"|reburn$Treatment == "bn"),]

#Checking Normality Assumption for each fuel type (all fuel types are normally distributed)
by(reburn$diff, reburn$fuel_type, shapiro.test)

#Live
live <- reburn[(reburn$fuel_type == "Live"),]

bartlett.test(diff ~ Treatment, live) # Checking Homogeneity of Variance (checks out)

live_aov <- aov(diff ~ Treatment, data=live)
summary(live_aov)

TukeyHSD(live_aov)

#Dead
dead <- reburn[(reburn$fuel_type == "Dead"),]

bartlett.test(diff ~ Treatment, dead) # Checking Homogeneity of Variance (checks out)

dead_aov <- aov(diff ~ Treatment, data=dead)
summary(dead_aov)

TukeyHSD(dead_aov)


#CWD
cwd <- reburn[(reburn$fuel_type == "CWD"),]

bartlett.test(diff ~ Treatment, cwd) # Checking Homogeneity of Variance (checks out)

cwd_aov <- aov(diff ~ Treatment, data=cwd)
summary(cwd_aov)

TukeyHSD(cwd_aov)


#FWD
fwd <- reburn[(reburn$fuel_type == "FWD"),]

bartlett.test(diff ~ Treatment, fwd) # Checking Homogeneity of Variance (checks out)

fwd_aov <- aov(diff ~ Treatment, data=fwd)
summary(fwd_aov)

TukeyHSD(fwd_aov)

#Litter
litter <- reburn[(reburn$fuel_type == "Litter"),]

bartlett.test(diff ~ Treatment, litter) # Checking Homogeneity of Variance (checks out)

litter_aov <- aov(diff ~ Treatment, data=litter)
summary(litter_aov)

TukeyHSD(litter_aov)


#Reburn Period(looking at ALL PLOTS)
diff <- read.csv("carbondiff.csv")
reburn <- diff[(diff$period == "reburn"),]

#Checking Normality Assumption for each fuel type (Dead and Live are not normally distributed)
by(reburn$diff, reburn$fuel_type, shapiro.test)

##Couldn't analyze live and dead due to negative and zero values

#CWD
cwd <- reburn[(reburn$fuel_type == "CWD"),]

bartlett.test(diff ~ Treatment, cwd) # Checking Homogeneity of Variance (checks out)

cwd_aov <- aov(diff ~ Treatment, data=cwd)
summary(cwd_aov)

TukeyHSD(cwd_aov)


#FWD
fwd <- reburn[(reburn$fuel_type == "FWD"),]

bartlett.test(diff ~ Treatment, fwd) # Checking Homogeneity of Variance (checks out)

fwd_aov <- aov(diff ~ Treatment, data=fwd)
summary(fwd_aov)

TukeyHSD(fwd_aov)

#Litter
litter <- reburn[(reburn$fuel_type == "Litter"),]

bartlett.test(diff ~ Treatment, litter) # Checking Homogeneity of Variance (checks out)

litter_aov <- aov(diff ~ Treatment, data=litter)
summary(litter_aov)

TukeyHSD(litter_aov)

```

###Change in Total Carbon Stock ANOVA (Before and After Second Entry Burn)
```{r}
carb <- read.csv("plotaverages.csv")

carb$year <- as.factor(carb$year)

temp <- aggregate(carbon ~ Plot + year, carb, sum)

temp <- temp[which(temp$year == "2017" | temp$year == "2018"),]

temp$Treatment <- substring(temp$Plot, 1, 2)

temp <- temp[which(temp$Treatment == "bc" | temp$Treatment == "bn" | temp$Treatment == "bs"),]

#Check for Normality
by(temp$carbon, temp$year, shapiro.test) #Meets the assumption of normality

#Check for homogeneity of variance
bartlett.test(carbon ~ Treatment, temp) #Meets the assumption

#ANOVA looking at Pre and Post Reburn Carbon Stocks by Treatment
reburn <-  aov(carbon ~ year * Treatment, temp)

summary(reburn)

TukeyHSD(reburn)
```


###ANOVA of Second Entry Burn Emissions
```{r}
emit <- read.csv("reburnemissions.csv")

#Check Normality (meets assumption)
shapiro.test(emit$emissions)

#Check Homogeneity of Variance (meets assumption)
bartlett.test(emissions ~ treatment, emit)

#ANOVA looking at emission differences by treatment
emitted <- aov(emissions ~ treatment, emit)
summary(emitted)

TukeyHSD(emitted)

#Average Emissions by Treatment
aggregate(emissions ~ treatment, emit, mean)
```



###Total Surface Fuel Differences by Year and Treatment 
```{r}
fuel <- read.csv("plotaverages.csv")

fuel$year <- as.factor(fuel$year)

fuel <- fuel[which(fuel$fuel_type != "Live" & fuel$fuel_type != "Dead"),]

fuel <- fuel[which(fuel$year != '2004'),]

temp <- aggregate(carbon ~ Plot + year, fuel, sum)

temp$Treatment <- substring(temp$Plot, 1, 2)

#Check for Normality
by(temp$carbon, temp$year, shapiro.test)

#Check for Homogenetiy of Variance
bartlett.test(carbon ~ year, temp)


#ANOVA of Total Surface Fuel Carbon by Year

yearaov <- aov(carbon ~ year+Treatment, temp)
summary(yearaov)
TukeyHSD(yearaov)

```

###Total Surface Fuel Differences by Year and Treatment using Gridpoint Data
```{r}
fuel <- read.csv("gridpointsurfacefuel.csv")

fuel$year <- as.factor(fuel$year)


#Check for Normality
by(temp$carbon, temp$year, shapiro.test)

#Check for Homogenetiy of Variance
bartlett.test(carbon ~ year, temp)


#ANOVA of Total Surface Fuel Carbon by Year

yearaov <- aov(carbon ~ year+Treatment, temp)
summary(yearaov)
TukeyHSD(yearaov)
```

###Live Carbon Stability using Large Tree Data from 2008, 2011 and 2017
```{r}
live <- read.csv("largetree_livecarbave.csv")

live$carbon_mg_ha <- ((live$carbon_mg_ha *12)/4)

live$year <- as.factor(live$year)

#BC
bc <- live[which(live$Treatment == 'bc'),]
shapiro.test(bc$carbon_mg_ha)

ggplot(bc, aes(x=year, y=carbon_mg_ha)) + geom_point()+
  geom_smooth(method="lm") 

bc.lm <- lm(carbon_mg_ha ~ year, data=bc)
bc_rsd <- sd(bc.lm$residuals)

#BN
bn <- live[which(live$Treatment == 'bn'),]
shapiro.test(bn$carbon_mg_ha)

ggplot(bn, aes(x=year, y=carbon_mg_ha)) + geom_point()+
  geom_smooth(method="lm") 

bn.lm <- lm(carbon_mg_ha ~ year, data=bn)
bn_rsd <- sd(bn.lm$residuals)


#BS
bs <- live[which(live$Treatment == 'bs'),]
shapiro.test(bs$carbon_mg_ha)

ggplot(bs, aes(x=year, y=carbon_mg_ha)) + geom_point()+
  geom_smooth(method="lm") 

bs.lm <- lm(carbon_mg_ha ~ year, data=bs)
bs_rsd <- sd(bs.lm$residuals)

#UC
uc <- live[which(live$Treatment == 'uc'),]
shapiro.test(uc$carbon_mg_ha)

ggplot(uc, aes(x=year, y=carbon_mg_ha)) + geom_point()+
  geom_smooth(method="lm") 

uc.lm <- lm(carbon_mg_ha ~ year, data=uc)
uc_rsd <- sd(uc.lm$residuals)

#US
us <- live[which(live$Treatment == 'us'),]
shapiro.test(us$carbon_mg_ha)

ggplot(us, aes(x=year, y=carbon_mg_ha)) + geom_point()+
  geom_smooth(method="lm") 

us.lm <- lm(carbon_mg_ha ~ year, data=us)

us_rsd <- sd(us.lm$residuals)

#UN
un <- live[which(live$Treatment == 'un'),]
shapiro.test(un$carbon_mg_ha)

ggplot(un, aes(x=year, y=carbon_mg_ha)) + geom_point()+
  geom_smooth(method="lm") 

un.lm <- lm(carbon_mg_ha ~ year, data=un)
un_rsd <- sd(un.lm$residuals)

ave_all <- aggregate(carbon_mg_ha ~ Plot, live, mean)

ave_all$carbon_mg_ha <- as.numeric(ave_all$carbon_mg_ha)

ave_all$Treatment <- substring(ave_all$Plot, 1, 2)

ave_all$rsd <- NA
ave_all[which(ave_all$Treatment == "bc"), 'rsd'] <- bc_rsd
ave_all[which(ave_all$Treatment == "bn"), 'rsd'] <- bn_rsd
ave_all[which(ave_all$Treatment == "bs"), 'rsd'] <- bs_rsd
ave_all[which(ave_all$Treatment == "uc"), 'rsd'] <- uc_rsd
ave_all[which(ave_all$Treatment == "un"), 'rsd'] <- un_rsd
ave_all[which(ave_all$Treatment == "us"), 'rsd'] <- us_rsd


ave_all$Stability_Treatment <- ave_all$carbon_mg_ha/ave_all$rsd

ave_all$order <- NA

ave_all[which(ave_all$Treatment == "un"), 'order'] <- 1
ave_all[which(ave_all$Treatment == "bn"), 'order'] <- 2
ave_all[which(ave_all$Treatment == "uc"), 'order'] <- 3
ave_all[which(ave_all$Treatment == "bc"), 'order'] <- 4
ave_all[which(ave_all$Treatment == "us"), 'order'] <- 5
ave_all[which(ave_all$Treatment == "bs"), 'order'] <- 6

ave_all$order <- as.factor(ave_all$order)

ave_all$Stability_Treatment <- as.numeric(ave_all$Stability_Treatment)
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}

stab <- ggplot(ave_all, aes(x=order, y=Stability_Treatment)) + geom_point(aes(fill=order),size = 3, shape=24, colour="black") + scale_fill_brewer(palette="Dark2")+
 scale_x_discrete(labels=addline_format(c("Control", "Burn Only","Understory Thin","Burn Understory Thin", "Overstory Thin", "Burn Overstory Thin")))+
   ylab("Large Tree Carbon Stability") 

stab <- stab + theme_bw()+
    theme(axis.title.x=element_blank())+ theme(legend.position = "none")


#Linear Regression of Pre Drought Carbon and Carbon Stability
carb <- read.csv("carbonstability.csv")

carb_stab <- lm(carbon_mg_ha_08 ~ Stability_Treatment, carb)

summary(carb_stab)

plot <- ggplot(carb, aes(x=carbon_mg_ha_08, y=Stability_Treatment)) + geom_point(size = 2.5, aes(shape=Treatment, col=Treatment))+
  geom_smooth(method = "lm", color="black") +
  xlab("2008 Large Tree Live Carbon (Mg/ha)")+
  ylab("Live Tree Carbon Stability")

plot


#2008 Large Tree Live Carbon is not a good predictor of the Carbon Stability Value
```

###2008 Large Tree Live Carbon and Temporal Standard Deviation
```{r}
plot <- read.csv("tempsd_08.csv")

summary(lm(sd ~ Carb_08, data=plot))


##2008 Large Tree Live Carbon is a good predictor of Temporal Standard Deviation (but this is obvious....more carbon means the deviation is likley to be larger)

```

